---
title: "R Notebook"
output: html_notebook
---


The original dataset contains 1000 entries with 20 categorial/symbolic attributes prepared by Prof. Hofmann. In this dataset, each entry represents a person who takes a credit by a bank. Each person is classified as good or bad credit risks according to the set of attributes. The link to the original dataset can be found below.

It is almost impossible to understand the original dataset due to its complicated system of categories and symbols. Thus, I wrote a small Python script to convert it into a readable CSV file. Several columns are simply ignored, because in my opinion either they are not important or their descriptions are obscure. The selected attributes are:

Age (numeric)
Sex (text: male, female)
Job (numeric: 0 - unskilled and non-resident, 1 - unskilled and resident, 2 - skilled, 3 - highly skilled)
Housing (text: own, rent, or free)
Saving accounts (text - little, moderate, quite rich, rich)
Checking account (numeric, in DM - Deutsch Mark)
Credit amount (numeric, in DM)
Duration (numeric, in month)
Purpose (text: car, furniture/equipment, radio/TV, domestic appliances, repairs, education, business, vacation/others)

```{r}
setwd("/Users/pierrot/Documents/german-credit-risk")
```


```{r}
#rm(list = ls())

data <- read.csv("german_credit_data.csv", sep=",",dec=".", h=T)
options(digits=2)
```

```{r}
data$Purpose_ <- "autre"
data$Purpose_[data$Purpose == "car"] <- "car"
data$Purpose_[data$Purpose == "furniture/equipment"] <- "furniture/equipment"
data$Purpose_[data$Purpose == "radio/TV"] <- "radio/TV"
```

```{r}

data$Job_<-ifelse(data$Job < 2, c("NoSkill"), c("Skill"))
```

```{r}

data$Housing_<-ifelse(data$Housing == "own", c("own"), c("NoOwn"))
```

```{r}

data$sex_num<-ifelse(data$Sex == "male", 1, 0)
data$job_num<-ifelse(data$Job_ == "Skill", 1, 0)
data$housing_num<-ifelse(data$Housing_ == "own", 1, 0)
```

PErmet de voir les premières lignes
```{r}
head(data)
```

Permet de voir le type des variables 
```{r}
str(data)
```

Permet d'avoir une vu d'ensemble
```{r}
summary(data)
```
`

Permet de voir les dernières lignes
```{r}
tail(data)
```
Proportion des variables catégorielles 
```{r}
prop.table(table(data$Job))
prop.table(table(data$Housing))
prop.table(table(data$Saving.accounts))
prop.table(table(data$Checking.account))
prop.table(table(data$Purpose))
prop.table(table(data$Sex))
```

Vérifier sil y a des valeurs vides
```{r}
length(which(is.na(data$Age)))
length(which(is.na(data$Sex)))
length(which(is.na(data$Job)))

length(which(is.na(data$Housing)))
length(which(is.na(data$Checking.account)))
length(which(is.na(data$Saving.accounts)))

length(which(is.na(data$Credit.amount)))
length(which(is.na(data$Duration)))
length(which(is.na(data$Purpose)))

```
Remplacer par la variables la plus frequentes
```{r}
data$Checking.account[is.na(data$Checking.account) == TRUE] <- 'little'
data$Saving.accounts[is.na(data$Saving.accounts) == TRUE] <- "quite rich"
```

```{r}
table(data$Sex,data$Job_)
```
```{r}
table(data$Sex,data$Housing)
```
```{r}
aggregate(data[, c(2,8:9)], list(data$Sex), mean)
```
```{r}
aggregate(data[, c(2,8:9)], list(data$Purpose_), mean)
```
```{r}
table(data$Sex,data$Purpose_)
```

```{r}
table(data$Job,data$Purpose)
tbl_purp_job <- table(data$Job,data$Purpose)
```

```{r}
df_tbl_purp_job <- as.data.frame(tbl_purp_job)
```


```{r}
table(data$Job)
```

Les personnes moins diplomés empruntent moins et sur une durée plus courte
```{r}
aggregate(data[, c(2,8:9)], list(data$Job_), mean)
```



```{r}
table(data$Purpose_)
```


```{r}
table(data$Job_)
```


```{r}
table(data$Job_,data$Purpose_)
```

```{r}
table(data$Sex,data$Housing)
```

```{r}
table(data$Purpose_,data$Housing_)
```


```{r}
install.packages("dplyr")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```

```{r}
install.packages("factoextra")
install.packages("NbClust")
```


```{r}
data_num <- data %>% 
  dplyr::select(Age, sex_num, job_num, housing_num, Credit.amount:Duration)
```

```{r}
data_num_sc <- scale(data_num)
```

```{r}
str(data_num_sc)
```
```{r}
km.out = kmeans(data_num_sc,centers=4,nstart =20)
```

```{r}
km.out$cluster
```
```{r}
km.out$tot.withins
```

```{r}
library(factoextra)
library(NbClust)
```
```{r}

# Elbow method
fviz_nbclust(data_num_sc, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")
```

```{r}
# Silhouette method
fviz_nbclust(data_num_sc, kmeans, method = "silhouette")+
labs(subtitle = "Silhouette method")
```
```{r}
# nboot = 50 to keep the function speedy.
# recommended value: nboot= 500 for your analysis.
# Use verbose = FALSE to hide computing progression.
set.seed(123)
fviz_nbclust(data_num_sc, kmeans, nstart = 25, method = "gap_stat", nboot = 50)+
labs(subtitle = "Gap statistic method")
```

```{r}
km.out2=kmeans(data_num_sc,centers=2,nstart =20)
pairs(data_num_sc, col=c(1:2)[km.out2$cluster])
```
```{r}
library(factoextra)
fviz_cluster(km.out2, data_num_sc, ellipse.type = "norm")
```

dashboard
encourt total
encourt moyen
purpose/encourt/duration
duration moyenne

segmentation
clustering